Figures to insert in text

### 2018-04-10 Chapters/Chapter02.tex
sdm re-org
recall theory intro

Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics. This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present. For instance, when one is walking on a dark alley and is afraid of something, one cannot explain where one's fear comes from. One just feels it. We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; a general lack of activity; that very feeling that something isn't right... without knowing precisely what isn't right, or even what `right' means... etc. --- recalling past experiences from memory and thus generating the feeling. Our memory is able to make a parallel between previous experiences and the clues. Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger. This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.

@@

\chapter{Results (i): Framework validation}
\input{./Chapters/sdm-validation.tex}

\chapter{Results (ii): Critical distance}
\input{./Chapters/sdm-critical-distance.tex}

@@

\chapter{Results (iv): Generalized read operation}

@@

\chapter{Results (v): Performance}
\input{./Chapters/sdm-performance.tex}

>>>>

Suddenly, you feel the danger.  You tighten your grip on the child's hand and step up your walking.  In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everthing you think of is to get out to safety.  You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exatcly is causing your fear. One just feels it. We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; a general lack of activity; that very feeling that something isn't right... without knowing precisely what isn't right, or even what `right' means... etc. --- recalling past experiences from memory and thus generating the feeling. Our memory is able to make a parallel between previous experiences and the clues. Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger. The whole process happens so fast it feels like it only took a single unit of psychological time.

Let us compare this to what a child might do.

Sparse Distributed Memory (SDM) \citep{Kanerva1988} is a mathematical model of long-term memory that has a number of neuroscientific and psychologically plausible dynamics. This model may be applied in all sort of applications because of its incredible ability to closely reflect the human capacity to remember past experiences from clues of the present.

This flexibility into mapping one situation in another is an essential human feature which is hard to replicate in computers.

\section{Desiderata for a theory of memory}

Psychological
miller magic number; when it instantly remembers; when it doesn't recall anything at all; when tip-of-the-tongue... how does this recall process work?

Neuroscientific (bioplausible)
neurons can die; what do neurons do?  how are things stored?  how are they recalled?

Computational (fully experimental, with no voodoo-like mysterious properties)

3 sparse distributed memory 11

3.1 Neurons as pointers 17

3.2 Concepts 17

3.3 Read operation 18

3.3.1 Generalized read operation

3.4 Critical Distance 20

4 framework architecture 23

4.1 Bitstring 24

4.2.2 OpenCL kernels

4.3 Counters 27

4.4 Read and write operations

5 results (i): performance

5.1 Kernels comparison 38

5.2 Scanners comparison 39

5.3 Read and write operations

6.1 Distance between random bitstrings

6.2 Number of activated hard locations

6.3 Intersection of two circles 63

6.4 Storage and retrieval of sequences

5.3.1 Summary of results

6 results (ii): framework validation

4.1.1 The distance between two bitstrings 24

4.2 Address space 25

4.2.1 Scanning for activated hard locations 25

6.4.1 k-fold memory using only one SDM

7 results (iii): loss of neurons 67

8 results (iv): critical distance 73

8.1 A deviation from the equator distance? 77

8.2 Counter bias 80

8.3 Read bias 83

8.4 Critical distance of 209 85

9 results (v): generalized read operation 87

10 results (vi): supervised classification
application

11 results (vii): image noise filtering application 101

12 results (viii): the possibility of unsupervised reinforcement learning 107

13 results (ix): information-theoretical write operation 115

@@

\chapter{Results (i): Performance}
\input{./Chapters/sdm-performance.tex}

\chapter{Results (ii): Framework validation}
\input{./Chapters/sdm-validation.tex}

@@

\chapter{Results (iv): Critical distance}
\input{./Chapters/sdm-critical-distance.tex}

\chapter{Results (v): Generalized read operation}

@@

%\chapter{Results (v): Performance}
%\input{./Chapters/sdm-performance.tex}

### 2018-04-10 Chapters/sdm-applications.tex
recall theory intro

The images had 31 pixels of width and 32 pixels of height, totaling 992 pixels per image. Each image was mapped into a 1,000-bit bitstring in which the bits were set according to the color of each pixel of the image. So, white pixels were equal to bit 0, and black pixels to bit 1. The eight remaining bits were all set to zero. This was a bijective mapping (or one-to-one mapping), i.e., there was only one bitstring for each image, and there was only one image for each bitstring.

@@

A total of 62 classification groups have been trained in the SDM. For each of them, it was generated a random bitstring. Thus, the groups' bitstrings were orthogonal between any two of them. There is one image for each of the 62 groups in Figure \ref{fig-classification-groups}. Notice that the SDM has never seen a single image with no noise.

@@

A final note is in order.  In no way is it claimed here that these bitstrings should form a plausible representation of letters in the human mind (See, for instance, \citet{hofstadter1995seeing} for a marvelous discussion of the subtleties and fluidity involved in that process).  These letters should merely be seen as \emph{invariant patterns} that are processed by the memory, which is able to capture their underlying invariant structure even when presented only under heavy noise.

>>>>

The images had 31 pixels of width and 32 pixels of height, totaling 992 pixels per image\footnote{A crucial note is in order.  While others have proposed similar models \citep{hong_character_1991}, in no way is it claimed here that these bitstrings should form a plausible representation of letters in the human mind (See, for instance, \citet{hofstadter1995seeing} for a marvelous discussion of the subtleties and fluidity involved in that process).  These letters should merely be seen as \emph{invariant patterns} that are processed by the memory, which is able to capture their underlying invariant structure even when presented only under heavy noise.  Recall that any rezising, any rotation, or any mere shifting of these letters will take them out of the critical distance for SDM reading.}. Each image was mapped into a 1,000-bit bitstring in which the bits were set according to the color of each pixel of the image. So, white pixels were equal to bit 0, and black pixels to bit 1. The eight remaining bits were all set to zero. This was a bijective mapping (or one-to-one mapping), i.e., there was only one bitstring for each image, and there was only one image for each bitstring.

@@

A total of 62 classification groups have been trained in the SDM. For each of them, it was generated a random bitstring. Thus, the groups' bitstrings were orthogonal between any two of them. There is one image for each of the 62 groups in Figure \ref{fig-classification-groups}. Notice that the memory has never seen a single image without noise.

@@
